\mythesistitleNOTSYNPOSIS

### 2012-01-12 src/030chap3.tex
Added more stuff to chapter 2. see diff.
partative set theory TBD in chapter 2. beginning to work on chapter 3 now.

\section{Introduction}
\label{sec:myresearchintro}
In Section~\ref{sec:motive} we see that consecutive-ones property and
its equivalent problem of interval labeling of a hypergraph is a
special case of the general problem of tree path labeling of path
hypergraphs.  The problem of consecutive-ones property testing can be
easily seen as a simple constraint satisfaction problem involving a
hypergraph or a system of sets from a universe. Every column (row) of
the binary matrix can be converted into a set of non-negative integers
which are the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to its
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the rows (columns). The cardinality of the interval
assigned to it must be same as the cardinality of the set, and the
intersection cardinality of any two sets must be same as the
intersection cardinality of the corresponding intervals. This is a
necessary and sufficient condition. \tnote{(ICPIA) put reference to
  survey chapter}

Naturally, intervals are paths from a tree with maximum degree
two. Thus the interval assignment problem can be generalized into path
assignment problem from any tree. We refer to this as the {\em tree
  path labeling problem}. To elaborate, the problem is defined as
follows -- given a hypergraph $\cF$ from universe (vertex set) $U$ and
a target tree $T$, does there exist a bijection from $U$ to the
vertices of $T$ such that when applied to elements of each hyperedge,
it gives a path on $T$.  More formally, the problem definition is as
defined by \CFTPL.

@@

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. Hence we consider two special cases of \CFTPL -- namely, \CFTPLINT
and \CFTPLKTREE on special target trees, intervals and \kstar,
respectively.  Section~\ref{sec:spltargettree} discusses these
problems.

@@

An algorithm for \CFTPL on general trees which is less efficient than
polynomial time is presented in
Section~\ref{sec:norestraint}. \tnote{put some more description
  here. refer walcom.}

@@

{\tt --- WALCOM}

@@

@@

% \begin{enumerate}[{(\icpplpr\ }i) \ \ \ ]
% \item \label{pr:i} $|S| = |\cl(S)|$ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $S \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ \ \ \ \
%   \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all distinct $S_1, S_2 \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
%   \cap \cl(S_3)|$ \ \ \ for all distinct $S_1, S_2, S_3 \in \cF$
% \end{enumerate}
\begin{enumerate}[{(\icpplpr\ }i)]
\item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
  % \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all distinct $S_1, S_2 \in \cF$
  % \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
  \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
\end{enumerate}

@@

{\tt WALCOM begin}

@@

\end{proof} {\tt WALCOM end}

@@

{\tt WALCOM begin}
\begin{lemma}
  \label{lem:invar3}
  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
  are valid.
  % Moreover, $X$ as defined in the algorithm is non-empty if this is
  % an ICPPA.
\end{lemma}
\begin{proof}
  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
  contained in at least two sets due to our assumption. Let $S_{i_2}$
  be a second set that contains $x$. We know $v \notin
  P_{i_2}$. Therefore there cannot exist a permutation that maps
  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
  that this is an ICPPA. Thus $X$ cannot be empty.

@@

\noindent
To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
$T$. After the leaves have been assigned to specific elements from
$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
original problem.%\qed
\end{proof}

{\tt WALCOM end}

@@

Consider a special case of ICPPL with the following properties.
\begin{enumerate}
\item Given tree $T$ is a path. Hence, all path labels are interval
  labels.
\item Only pairwise intersection cardinality preservation is
  sufficient. i.e. property (iii) in ICPPL is not enforced.
\item The filter algorithms do not have {\em \bf exit} statements.
\end{enumerate}
This is called an Intersection Cardinality Preservation Interval
Assignment (ICPIA) \cite{nsnrs09}. This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

@@

\noindent
Suppose the given hypergraph has a feasible path labeling to the given
tree $T$. Let this labeling be $\cl$. Now we present an algorithm that
lists the sequence of edges that maintain a set of properties
necessary for them to have a feasible tree path labeling.
\begin{enumerate}[I.]
\item {\bf Step 1}
  \begin{enumerate}
  \item {\bf {There are type 1 edges, $\mathbf{|\cT_1^i| > 0}$}:}
    % , and for each such $X$ assign the path in $R_i$ of length
    % $s(X)$ starting at $v_{i,k+2}=r$.  $X$ is then removed from
    % $\cO_i$, and $p_2(X)=s(X)$, and $s(X)=0$.If any intersection
    % cardinality is violated because of this assignment, EXIT
    % reporting the non-existence of a feasible assignment.
    {If there is only one $\cT_1^i$ hyperedge, label it to the unique
      path in $R_i$ of length $s(X)$ starting at $v_{i(k+2)}=r$. If
      there are more than one $\cT_1^i$ hyperedges, for any pair of
      hyperedges $S_1, S_2 \in \cT_1^i$, one of the following is
      true. Let $\cl'(X) \subseteq R_j$ denote the subpath of length
      $p_1(X)$ that has been assigned to $X$ from a previously
      considered ray $R_j$.}
    \begin{enumerate}[{\em Case }1. ]
    \item $|\cl'(S_1) \cap \cl'(S_2)| = 1$ and $|S_1 \cap S_2| > 1$
      i.e. they are in different rays and their residues must contain
      their intersection. Hence assign both to $R_i$.
    \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| >
      |\cl'(S_1) \cap \cl'(S_2)|$ i.e. they are in the same rays and
      their residues must contain their intersection. Hence assign
      both to $R_i$.
    \item $|\cl'(S_1) \cap \cl'(S_2)| > 1$ and $|S_1 \cap S_2| =
      |\cl'(S_1) \cap \cl'(S_2)|$. i.e. the intersection cardinality
      has been met. Hence assign only $S_1$ to $R_i$ and $S_2$ must be
      assigned to some $i \ge p$.
    \end{enumerate}
    Remove the labeled edge or edges from $\cO_i$ and add it to
    $\cO_{i+1}$. Also do not add it to $\cT_1^{i+1}$ and add to
    $\cL_2^{i+1}$.
  \item {\bf {There are no type-1 edges nor} super-marginal type-2
      hyperedges:} If all type-2 edges of length at most $k+1$, EXIT
    by saying there is no ICPPL.  Otherwise, Find an $X$ of type-2
    such that $|X| \geq k+2$, assign it the unique path starting at
    $v_{i(k+1)}=r$ of length $k+2$, and set $p_1(X)=k+2,
    s(X)=|X|-(k+2)$, assign it the unique path starting at
    $v_{i(k+1)}=r$ of length $k+2$, mark this as a type-1 hyperedge,
    and add it to $\cO_{i+1}$ after removing it from $\cO_i$.
    % \remove{Next, iteratively, consider each type-1 edge $X \in
    %   \cO_i$, and if it intersects a previously assigned $Y$ such
    %   that
    %   $\cl(Y) \subseteq R_i$, assign a path of length $s(X)$
    %   containing $v_{i(k+2)}=r$, set $p_2(X)=s(X), s(X)=0$, and
    %   remove
    %   $X$ from $\cO_i$.}\tnote{covered in step 2}

  \item {\bf There are super-marginal type-2 hyperedges {and no type-1
        edges:}} Let $X$ be a super-marginal hyperedge. If $|X| \leq
    k+2$, then $X$ is assigned the path of length $|X|$ starting at
    $v_{i1}$, and $p_1(X) = |X|, p_2(X)=0$. $X$ is removed from
    $\cO_i$ and not added to $\cO_{i+1}$.  If $|X| > k+2$, then
    $p_1(X)=k+2$, assign it the unique path starting at $v_{i(k+1)}=r$
    of length $k+2$ and $s(X) = |X| - (k+2)$.  In this case, $X$ is
    classified as a Type-1 edge, removed from $\cO_i$, and added to
    $\cO_{i+1}$.
  \end{enumerate}
\item {\bf Step 2:} Iteratively, a hyperedge $X$ is selected from
  $\cO_i$ that has an overlap with one of the hyperedges $Y$ such that
  $\cl(Y) \subseteq R_i$, and a unique path is assigned to $X$.  The
  path, say $U(X) \subseteq R_i$, that is assigned can be decided
  unambiguously since the $X \overlap Y$, and all intersection
  cardinalities can be preserved only at one of the ends of $\cl(Y)$.

  \noindent
  Let $\cl(X)$ denote the unique path assigned to $X$.  If $X$ is a
  type-2 hyperedge and if the unique path of length $|X|$ does not
  contain $r$, then $p_1(X) = |X|, p_2(X)=0$, and $X$ is removed from
  $\cO_i$.  In the case when $\cl(X)$ has to contain $r$, then
  $p_1(X)$ is the length of the path, $p_2(X)=0$, and $s(X) =
  |X|-p_1(X)$. Further $X$ is classified as a type-1 hyperedge, added
  to $\cO_{i+1}$ and removed from $\cO_i$.  In the case when $X$ is a
  type-1 hyperedge, then we check if $U(X)$, which is of length $s(X)$
  contains $r$. If it does, then we assign $\cl(X) \leftarrow \cl(X)
  \cup U(X)$, remove $X$ from $\cO_i$ and do not add it to
  $\cO_{i+1}$.  If not, then we {\bf Exit} reporting that an
  assignment cannot be found.  The iteration ends when no hyperedge in
  $\cO_i$ has an overlap with a hyperedge assigned to $R_i$.
\end{enumerate}

\noindent
The order in which they are assigned is the sequence of hyperedges
with the following properties. Note that if such a sequence cannot be
computed, $\cO$ cannot have a feasible path labeling from $T$.

\noindent
\tnote{need to formalize the below properties} In the following lemmas
we identify a set of necessary conditions for $\cF$ to have an ICPPL
in the $k$-subdivided star $T$.  If during the execution of the above
algorithm, one of these necessary conditions is not satisfied, the
algorithm exits reporting the non-existence of an ICPPL.
\begin{lemma}
  Let all hyperedges in $\cO_i$ be type-1 edges.  Then there is a
  maximal subset $\ccT_i \subseteq \cO_i$ with the following
  properties:
  \begin{enumerate}
  \item $\ccT_i$ form an inclusion chain.
  \item For all $X \in \ccT_i$, $s(X) \leq k+2$, and There is a an $X
    \in \ccT_i$ such that $s(X)=k+2$.
  \end{enumerate}
\end{lemma}
\begin{lemma}
  If there are no \marginal type-2 edges in $\cO_i$, then there exists
  at least $p-i$ type-2 hyperedges $X \in \cO_i$ such that $|X| \geq
  k+2$.
  % is a exist type-2 hyperedges of length at least $k+2$, and there
  % must be at least $r-i$ such type-2 hyperedge, basically at least
  % one per ray.  Each hyperedge contained in such an hyperedges of
  % length at least $k+2$, is transitively related under the overlap
  % relation to one that gets assigned a path that contains
  % $r$. Basically, the proof will argue that otherwise, there will be
  % a \marginal type-2 hyperedge, or that it is infeasible.  Also, it
  % is here we assume that we assume that the size of the two
  % universes is the same.
\end{lemma}
\begin{lemma}
  At the end of Step-1 in the $i$-th iteration, if one hyperedge $X$
  of type-2 is such that $\cl(X) \subseteq R_i$, then all other
  hyperedges in $\cO_i$ are connected to $X$ in the overlap component.
\end{lemma}
\begin{lemma}
  At the end of Step-2, If control has exit at any time, there is no
  ICPPL. If control has not exit, then $R_i$ is saturated.  No
  hyperedge of $\cO_{i+1}$ will get a path from $R_i$ in the future
  iterations.  No type-2 hyperedge of $\cO_{i+1}$ will get a path from
  $R_i$.  Basically $R_i$ is done.
\end{lemma}
\begin{lemma}
  Finally, we need to prove that the assignment is an ICPPL. Secondly,
  if there is a permutation then maps sets to paths, then it is indeed
  an ICPPL, and our algorithm will basically find it.  It is a unique
  assignment upto permutation of the leaves.
\end{lemma}

\begin{lemma}
  \label{lem:sup-mar}
  If $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible tree
  path labeling to tree $T$, then $\cl(X)$ will
  contain a leaf in $T$.\\
  On the otherhand, if $\cl(X)$ has a leaf in $T$, $X$ is marginal but
  may not be super-marginal.
\end{lemma}
\begin{proof}
  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
  Since $X$ is in a connected overlap component, there exists $Y_1 \in
  \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
  at least one vertex closer to the leaf in $R_i$ than any vertex in
  $X$. Similarly with the same argument there exists $Y_2 \in \cF$
  with same subset and overlap relation with $X$ except it has has at
  least one vertex farther away from the leaf in $R_i$ than any vertex
  in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
  inclusion chain which contradicts that assumption $X$ is
  super-marginal. Thus the claim is proven.%\qed
\end{proof}

\noindent
Consider the overlap graph $\bO(\cF)$ of the given hypergraph
$\cF$. Let $S_{sm} \in \cF$ be such that it is a super-marginal
hyperedge.  Algorithm~\ref{al:ktree-label} uses $S_{sm}$ along with
the overlap graph $\bO(\cF)$ to calculate the feasible tree path
labeling to the $k$-subdivded tree $T$.

\begin{algorithm}[h]
  \caption{{\tt compute-ksubtree-path-labeling($X, \cF, T$)}}
  \label{al:ktree-label}
  \begin{algorithmic}[\lndisplay]
    \IF{$X = S_{sm}$}
    \STATE -- TBD --\\
    \ELSE
    \STATE -- TBD --\\
    \ENDIF

  \end{algorithmic}
\end{algorithm}

>>>>

\section{Introduction}
\label{sec:myresearchintro}
In Section~\ref{sec:motive} we see that consecutive-ones property and
its equivalent problem of interval labeling of a hypergraph is a
special case of the general problem of tree path labeling of path
hypergraphs.  The problem of consecutive-ones property testing can be
easily seen as a simple constraint satisfaction problem involving a
hypergraph or a system of sets from a universe. Every column (row) of
the binary matrix can be converted into a set of non-negative integers
which are the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to an
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the universe. The cardinality of the interval
assigned to it must be same as the cardinality of the set, and the
intersection cardinality of any two sets must be same as the
intersection cardinality of the corresponding intervals. This is a
necessary and sufficient condition. \tnote{(ICPIA) put reference to
  survey chapter}

Naturally, intervals are paths from a tree with maximum degree
two. Thus the interval assignment problem can be generalized into path
assignment problem from any tree. We refer to this as the {\em tree
  path labeling problem of path hypergraphs}. This is analogous to the
interval labeling problem in literature \cite{kklv10} to interval
hypergraphs. To elaborate, the problem is defined as follows -- given
a hypergraph $\cF$ from universe (hypergraph vertex set) $U$ and a target tree
$T$, does there exist a bijection $\phi$ from $U$ to the vertices of $T$ such
that for each hyperedge when applied to its elements, $\phi$ gives a path on
$T$.  More formally, the problem definition is as defined by \CFTPL.

@@

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. Hence we consider two special cases of \CFTPL\ -- namely, \CFTPLINT
and \CFTPLKTREE on special target trees, intervals and \kstars,
respectively.  Section~\ref{sec:spltargettree} discusses these
problems.

@@

An algorithm for \CFTPL on general trees which is less efficient than
polynomial time is presented in
Section~\ref{sec:norestraint}. \tnote{put some more description
  here. refer WG11.}

@@

\temptext{ --- WG11 begin --- }

@@

\temptext{ --- WG11 end --- }

@@

\begin{enumerate}[{(\icpplpr\ }i)]
\item \label{pr:i} $|S| = |\cl(S)|$ for all $S \in \cF$
  % \vspace{\topshrink}
\item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$ for all distinct $S_1, S_2 \in \cF$
  % \vspace{\topshrink}
\item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
  \cap \cl(S_3)|$ for all distinct $S_1, S_2, S_3 \in \cF$
\end{enumerate}

@@

\temptext{ --- WG11 begin --- }

@@

\end{proof} 
\temptext{ --- WG11 end --- }

@@

\temptext{ --- WG11 begin --- }
\begin{lemma}
  \label{lem:invar3}
  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
  are valid.
  % Moreover, $X$ as defined in the algorithm is non-empty if this is
  % an ICPPA.
\end{lemma}
\begin{proof}
  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
  contained in at least two sets due to our assumption. Let $S_{i_2}$
  be a second set that contains $x$. We know $v \notin
  P_{i_2}$. Therefore there cannot exist a permutation that maps
  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
  that this is an ICPPA. Thus $X$ cannot be empty.

@@

\noindent
To summarize, run algorithm \ref{perms} and \ref{leafasgn} on
$T$. After the leaves have been assigned to specific elements from
$U$, remove all leaves from $T$ to get new tree $T_0$. The leaf
assignments are in $\cL_0$. Since only leaves were removed $T_0$ is
indeed a tree. Repeat the algorithms on $T_0$ to get leaf assignments
$\cL_{1}$. Remove the leaves in $T_0$ to get $T_1$ and so on until the
pruned tree $T_d$ is a single path. Now run ICPIA algorithm on $T_d$
to get permutation $\sigma'$. The relation $\cL_0 \cup \cL_1 \cup
.. \cup \cL_{d} \cup \sigma'$ gives the bijection required in the
original problem.%\qed
\end{proof}
\temptext{ --- WG11 end --- }

@@

Consider a special case of ICPPL with the following properties when the tree $T$
is a path.  Hence, all path labels are can be viewed as intervals assigned to the
sets in $\cF$.  It is shown, in \cite{nsnrs09}, that the filtering algorithms outlined
above need only preserve pairwise intersection cardinalities, and higher level
intersection cardinalities are preserved by the Helly Property of intervals.  Consequently,
the filter algorithms do not need to ever evaluate the additional check to {\em \bf exit}.
%\begin{enumerate}
%\item Given tree $T$ is a path. Hence, all path labels are interval labels.
%\item Only pairwise intersection cardinality
%  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
%\item The filter algorithms do not have {\em \bf exit} statements.
%\end{enumerate}
%This is called an Intersection Cardinality Preservation Interval
%Assignment (ICPIA) \cite{nsnrs09}. 
This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

@@

\temptext{repetition}

In this section we consider the problem of assigning paths from a
$k$-subdivided star $T$ to a given set system $\cF$ such that each set
$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
results only for the case when overlap graph $\bO(\cF)$ is connected.
%The
%overlap graph is well-known from the work of
%\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
%\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
%$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
%have a non-empty intersection but neither of them is contained in the
%other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
%correspond to the sets in $\cF$, and the vertices corresponding to the
%hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
%that the intersection graph of $\cF$, $\bI(\cF)$ is different from
%$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
A connected component of $\bO(\cF)$ is called an overlap component of
$\cF$.  An interesting property of the overlap components is that any
two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
$\cO_2$.  This containment relation naturally determines a
decomposition of the overlap components into rooted containment trees.
We consider the case when there is only one rooted containment tree,
and we first present our algorithm when $\bO(\cF)$ is connected.  It
is easy to see that once the path labeling to the overlap component in
the root of the containment tree is achieved, the path labeling to the
other overlap components in the rooted containment tree is essentially
finding a path labeling when the target tree is a path: each target
path is a path that is allocated to sets in the root overlap
component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
connected graph. Recall that we also consider the special case when
all hyperedges are of cardinality at most $k+2$.  By definition, a
$k$-subdivided star has a central vertex which we call the {\em root},
and each root to leaf path is called a {\em ray}.  First, we observe
that by removing the root $r$ from $T$, we get a collection of $p$
vertex disjoint paths of length $k+1$, $p$ being the number of leaves
in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
all $R_i$.

\subsection{Description of the Algorithm}
In this section the given hypergraph $\cF$, the $k$-subdivided star
and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
respectively.  In particular, note that the vertices of $\cO$
correspond to the sets in $\cF$, and the edges correspond to the
overlap relation.

\noindent
For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
 numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
 $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
 $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
 the lengths of subpaths of the path assigned to $X$ from at most two
 rays. We also maintain another parameter called the {\em residue} of
 $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
 length that must be assigned to $X$ which must be from another
 ray. For instance, if $X$ is labeled a path from only one ray, then
 $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.

 \xnoindent The algorithm proceeds in iterations, and in the $i$-th
 iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
 hyperedge that has been assigned a path is considered.  At the
 beginning of each iteration hyperedges of $\cO$ are classifed into
 the following disjoint sets.
 \begin{enumerate}
 \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
   labeled with a path which does not contain $r$ in one of the
   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
 \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
   with two subpaths of $\cl(X)$ containing $r$ from two different rays
   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
     p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
     been labeled with one path containing $r$ from a single ray in one
     of the previous iterations. Here, $p_1(X)$ denotes the length of
     the subpath of $\cl(X)$ that $X$ has been so far labeled
     with. \\
     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
       \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
     labeled with a path in any previous iteration.\\
     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
       X \in \cO}$
 \end{enumerate}
% \vspace{-2mm}
% \begin{align*}
%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
% \end{align*}

\noindent
The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
the following rules. Also the end of the iteration, $\cL_1^{i+1},
\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
case-specific changes mentioned in the rules below.

\noindent
\begin{enumerate}[I.]
\item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
  super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
  exit reporting failure.  Else, assign to each $X_j \in S$, the path
  from $R_j$ such that the path contains the leaf in $R_j$.  This path
  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
  Hyperedges in $S$ are not added to $\cO_2$ but are added to
  $\cL_1^2$ and all other hyperedges are added to
  $\cO_2$.
\item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
  that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
  Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
  If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
  any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
  $X$ has a partial path assignment, $\cl'(X)$ from a previous
  iteration, say from ray $R_j$. Then, $Y$ is
  chosen such that $X \cap Y$ has a non-empty intersection with a ray
  different from $R_j$.  The key things that are done in assigning a
  path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
  $\cl(X)$ would overlap is found, and then based on this the
  existence of a feasible assignment is decided.  It is important to
  note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
  feasible assignment.  Therefore, the notion of the {\em end} at
  which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
  path, there are two end points.
  \begin{enumerate}
  \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
      depends on $X \cap Z$:} If $X \cap Z
    \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
    end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
    \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
    that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
  \item {\em Any path of length $s(X)$ at the appropriate end contains
      $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
    as in step~\ref{iendpoint} this the unique path of length $s(X)$
    should end at $r$.  If not, we exit reporting failure.  Else,
    $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
    three-way intersection cardinality is violated with this new
    assignment, then exit, reporting failure.  Otherwise, $X$ is added
    to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
    after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
    the root and including it. Clearly, the vertices $\cl(X)$ or
    $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
    three way intersection cardinality is violated due to this
    assignment, exit, reporting failure.  Otherwise,
    $p_1(X)$ is updated as the length of the assigned path, and $s(X)
    = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
    If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
  \item {\em The unique path of length $s(X)$ overlapping at the
      appropriate end of $Y$ does not contain $r$:} In this case,
    $\cl(X)$ is updated to include this path.  If any three way
    intersection cardinality is violated, exit, reporting failure.
    Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
    added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
  \end{enumerate}
\end{enumerate}

\noindent {\bf Proof of Correctness and Analysis of Running Time:} It
is clear that the algorithm runs in polynomial time, as at each step,
at most three-way intersection cardinalities need to be checked.
Further, finding super-marginal hyperedges can also be done in
polynomial time, as it involves considering the overlap regions and
checking if the inclusion partial order contains a single minimal
element.  In particular, once the super-marginal edges are identified,
each iteration involes finding the next hyperedge to consider, and
testing for a path to that hyperedge.  To identify the next hyperedge
to consider, we consider the breadth first layering of the hyperedges
with the zeroeth layer consisting of the super-marginal hyperedges.
Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
be considered by the algorithm.  Once a hyperedge is considered, the
path to be assigned to it can also be computed in constant time.  In
particular, in the algorithm the path to be assigned to $X$ depends on
$\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
edges are identified, the running time of the algorithm is linear in
the size of the input.  By the technique used for constructing prime
matrices \cite{wlh02}, the super-marginal edges can be found in linear
time in the input size.  Therefore, the algorithm can be implemented
to run in linear time in the input size.
The proof of correctness uses the following main properties:
\begin{enumerate}
\item The $k$-subdivided star has a very symmetric structure.  This
  symmetry is quantified based on the following observation -- either
  there are no feasible path labelings of $\cO$ using paths from $T$,
  or there are exactly $p!$ feasible path labelings.  In other words,
  there is either no feasible assignment, or effectively a unique
  assignment modulo symmetry.
\item The $p$ super-marginal hyperedges, if they exist, will each be
  assigned a path from distinct rays, and each such path contains the
  leaf.
\item For a candidate hyperedge $X$, the partial path assignment
  $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
  intersection with $\cl(Z)$.
\end{enumerate}
These properties are formalized as follows:
\begin{lemma}
  \label{lem:sup-mar}
  If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
  labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
  Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
  path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
  $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
  Since $X$ is in a connected overlap component, there exists $Y_1 \in
